Multi-stream portrait of the Cosmic web

### 2018-06-12 Draft/Chapter3/Paper2015.tex
spell checked

The problem of objective identification and quantitative characterization of anisotropic structures 
in the distribution of galaxies in space emerged after the first evidences of their existence
(see the review by \citealt{Oort1983} and the references therein). The first theoretical model predicting highly anisotropic concentrations in the mass distribution coming into existence at the non-linear stage of gravitational instability is known as the Zeldovich Approximation (the ZA) (\citealt{Zeldovich1970}, for further developments see also \citealt{Shandarin1989} and the references therein). The ZA predicted the formation of  `pancakes'  also known as the walls in the currently popular jargon. The later development
of the model by \citet{Arnold1982} predicted the formation of filaments along with the pancakes. \citet{Klypin1983a} and \citet{Shandarin1984} demonstrated that the filaments emerge in the cosmological N-body simulation in three-dimensional space.
However, they failed to identify the pancakes at $z=0$. Both the existence of filaments connecting compact clumps of matter and absence of pancakes were confirmed by \citet{Frenk1983}. Puzzled by the absence of the pancakes \citet{Klypin1983a}
speculated that insufficient mass resolution of the simulation was the cause  of the negative outcome. 
This has been unambiguously confirmed by recent simulations using a better numerical technique of computing a density field from the particle
coordinates in cosmological N-body simulations (~\citealt{Shandarin2012} and ~\citealt{Abel2012}). \citealt{Klypin1983a} 
also stressed that the most of filaments are  incorporated in 
`a single three-dimensional web structure'\footnote{The term `cosmic web' was coined by  \citet{Bond1996}.}. 
They admitted that
their simulation did not allow them to confirm the existence of pancakes between the filaments predicted by the ZA \citep{Arnold1982}.

Although the four archetypical elements of the cosmic web:  voids, walls/pancakes, filaments and haloes 
were predicted by ZA and confirmed in cosmological N-body simulation their identification and quantitative characterization
remains under vigorous debate
(see e.g. \citealt{Colberg2008}, \citealt{Elahi2013}, \citealt{Knebe2013}, \citealt{Hoffmann2014}).
The dark matter haloes are arguably the easiest objects to identify in N-body simulations. They can also be reliably associated with
observed objects like galaxies and clusters of galaxies.  But even in this case \citet{Knebe2013} refer to almost thirty different
halo finders suggested after 2000. Identifying filaments and pancakes/walls is far more controversial in both N-body simulations 
and galaxy catalogues. For instance, even estimating the global parameters of the  web  in N-body simulation such as the fractions of volume and mass
in voids, walls/pancakes, filaments and haloes produced quite different results. The estimates of volume 
fractions of voids range from 13  to 86\%  (\citealt{Cautun2014a}, \citealt{Falck2015}, \citealt{Forero-Romero2009a} , \citealt{Hahn2007} , \citealt{Aragon-Calvo2010a}). Similar estimates for walls/pancakes, filaments 
and haloes  are respectively 5-56\% 2-26\% and 0.1-1\%.  Estimates of the mass content vary in large ranges as well.

Large differences in the estimates of volume and mass fractions made by different groups are not surprising if we recognize
considerable differences in the definitions of the components of the cosmic web and numerical methods used in the estimates. Without trying to provide an exhaustive review of all definitions and techniques used for the quantitative morphological 
analysis of the web  we just briefly describe a few approaches in order to illustrate how different they could be. 
Some groups study the web morphology using only coordinates of simulation particles, while others use the particle velocities too. 
Transforming data from  point sets to the density and other fields on a grid is also often used
because fields allow to use a variety of mathematical techniques not available for the particle sets.
However this step can be done by a variety of methods some as simple as cloud-in-cell (CIC), or more complicated as smoothed-particle hydrodynamics (SPH), 
or using Voronoi and Delaunay tessellations
as in the Delaunay Tessellation Field Estimator (DTFE) method (~\citealt{Weygaert2009}, \citealt{Cautun2014a}).  
Recently a new method called a discrete persistent structure extractor (DisPerSE \citealt{Sousbie2011f},  \citealt{Sousbie2011e} allowing to identify haloes and other components of the web directly from the particles 
has been designed.  This method can be applied to the galaxy catalogues, for instance \citet{Sousbie2011e} applied it to SDSS catalogue and extracted the filaments (which are made available online).

An obvious advantage of methods based on particle coordinates, both based on the density field and directly on particle coordinates, is their applicability to redshift catalogues. 
The redshift catalogues like SDSS and 2dF provide only two angular coordinates and distances in redshift space. 
But cosmological N-body dark matter simulations provide the full dynamical information in six-dimensional
phase space. This additional information is very valuable providing a greater opportunity for understanding the 
physics of the web and developing a better theory of the web.

@@

Using the full six-dimensional information allows one to generate new fields which  provide 
additional useful information about the evolution and morphology of the  web. 
One of them is a multi-stream field in Eulerian space, which will be the focus of this chapter. 
Another example is the flip-flop field in Lagrangian space. In cosmological context it was first used in the ZA.
\citet{Vogelsberger2011} used it in a study of multi-stream structure of galaxy size haloes.
\citet{Shandarin2014a} applied it for identifying subhaloes  in dark matter haloes.
A similar although somewhat simplistic realization of this idea has been revealed in the ORIGAMI method
used for the analysis of the web ( \citealt{Falck2012},  \citealt{Falck2015} ).
 Although these fields cannot be used directly on observational data because the full phase-space information
 is not available, they provide much deeper insight into non-linear clustering of collision-less dark matter
and reveal new features of the web.

@@

In order to compute the multi-stream field we will use the tessellation scheme described in
 \citet{Shandarin2012} which is  also briefly discussed in Section \ref{subsec:method}.
  Using this methodology on the entire simulation box, we discuss the global behaviour of the number of streams in the cosmic web in Section \ref{sec:global}. 
  The tessellation technique we have utilized can be used to find multi-stream fields in smaller Eulerian boxes with very high resolution too. In Section \ref{sec:local} we study the local behaviour of multi-streams flows in regions around haloes the detected using friends-of-friends (FOF) technique.

@@

We have utilized the data from cosmological N-body simulations by Gadget-2 \citep{Springel2005c} for 100 $h^{-1}$ Mpc  and 200 $h^{-1}$ Mpc box 
sizes with $128^3$, $256^3$ and $512^3$ grids. Each particle is between  $10^9 - 10^{12} M_{\odot}$. 
The initial conditions and cosmological parameters are consistent with the Planck cosmology. We utilize the initial Lagrangian box and do a three-dimensional mapping onto corresponding evolved simulations. In addition, for local multi-stream analyses around haloes, we have utilized halo catalogues for each of these simulation boxes. These haloes are detected using FOF method considering objects with more than 20 particles found at linking length, $ b= 0.2$.  
%-------------------------------------------------------------------------------------------------------------------------------
\section{Multi-stream field calculation}
\label{subsec:method}

Phase space tessellation considers the dynamics of the particles similar to that of a standard N-body code. 
However the particles are nodes of the tessellation, and are just massless tracers of the flow.
Assuming that the uniform state is modelled by a simple rectangular grid, the particles are the nodes of the grid.
Each elementary cube of the grid is tessellated by five tetrahedra (\citealt{Shandarin2012}\footnote{For the description of an alternative type of the tessellation see \citet{Abel2012}.}) of which the vertices are the
vertices of the cube. Mass is assumed to be uniformly distributed within each tetrahedron and the tessellation remain
intact at all times. The tetrahedra of the tessellation change their shapes and volumes, the latter are used for
computing the densities of the tetrahedra. Despite the complicated deformations experienced by the three-dimensional
sub-manifold tessellated by the tetrahedra, it remains continuous. Projected on three-dimensional configuration space, the tetrahedra
may form complicated structures. The number of streams at a chosen point {\bf x} is simply the number of tetrahedra that contain the point. The diagnostic points are computationally convenient to choose on a regular grid which can be significantly finer than the original grid in Lagrangian space. The ratio of separation of particles on the initial unperturbed grid to the separation distance of points in the diagnostic grid  $l_{\rm part}/l_{\rm dg} $ will be referred to as the refinement factor in the rest of the chapter.

@@

Whereas for non-linear structures, our parameter space has more freedom in terms of number of streams. 
Similar to the density threshold, the number of streams - used as a local parameter - cannot distinguish unambiguously whether a point is in a wall, 
filament or halo. Only some parts of walls where there are only three streams can be identified locally without confusion. This is because the formation of a filament
requires at least five streams. A flip-flop along one axis  would produce a three-stream region which may be only a pancake. 
 Therefore  another flip-flop along the other axis in one of the streams
from previous stage is required to transform it into a three-stream flow. Thus the total becomes five. However, if the second flip-flop happens along the same axis 
the resultant structure will remain a wall. Therefore some points in the five-stream flows can be within walls while the other in filaments. The present simulations have no information about the evolution of the flip-flop field
therefore we rely on visual impressions initially to understand the transformation of walls into filaments and parts of filaments into  haloes. By inspection, we have identified all the regions with three streams as walls. Unfortunately, walls are difficult to display on paper since they essentially block the view in two-dimensional projection. Nevertheless, we have demonstrated and analysed walls on a smaller Eulerian box around haloes in Section \ref{sec:local} using a simple and reasonably effective approximation.

@@

The filamentary structure of regions with 17 or more streams (denoted as 17+ ) is shown for a slice of simulation box of size $100 h^{-1}$  Mpc and $128^3$  particles in Fig. \ref{fig:full}. It has to be noted that all the regions with 17+ streams are regions with 3+ streams. Thus, the filaments are just interior
parts  of walls with higher $n_{str}$. These are visually observed mostly at the intersections of walls. Further, at the intersections of multiple filaments, there are regions with locally maximum number of streams, signifying the most dense regions in the simulations i.e. the dark matter haloes as Fig. \ref{fig:4view}
illustrates.
By superimposing the positions from the FOF-halo catalogue, it is visually confirmed that the FOF haloes reasonably coincide with these high-streaming intersections, as Fig. \ref{fig:full} illustrates.

@@

One of the advantages of using tessellation is the freedom to compute densities at very high resolutions (\cite{Abel2012}, \cite{Shandarin2012}). We remind that the parameter `refinement factor' denotes the ratio of separation of the particles to the separation distance of points in the diagnostic grid as defined in Sec. 3. High refinement factors are extensively used in understanding stream behaviour not only in the halo environment, but inside the halo too (Section ~\ref{sec:local}). The volume fractions of resulting number of streams are
%normalized by multi-stream field resolution (i.e. dividing by the cube of refinement factor) are %
similar for all refinement factors as
shown in bottom of the Fig. ~\ref{fig:Vfr_all} and in Table ~\ref{tab:Compare_ref}. % Hence our triangulation algorithm is universal in the sense that it is independent of the refinement factors used in tessellating the box.   
Multi-stream fields calculated on low refinement factors are more noisy at high number of streams.

With same refinement factors, the mass fractions exhibit similar pattern for same $L/N$ as well (Fig.~\ref{fig:Mfr_all}). The simulation box with highest inter-particle distance (thus least mass resolution) has more mass particles in single streaming region, as tabulated in Table~\ref{tab:Compare_LN}, but decreases steeply thereafter (Table ~\ref{tab:Compare_Slopes}). Unlike the volume fraction, the behaviour of mass fraction has a systematic variation across different refinement factors. The mass fractions given in Table ~\ref{tab:Compare_ref} show that the single-streaming regions in the multi-stream fields with refinement factors 1 and 2 have higher mass fraction than in the fields with refinement factor of 8. Decreasing the resolution from refinement factor of 8 to 1 effectively introduces smoothing of the structure. This results in growth of mass fraction in voids ($n_{str} = 1$) and decreasing it in the web ($n_{str} > 1$). The multi-stream field is more robust as one can see in Fig.~\ref{fig:Vfr_all}. and \ref{fig:Mfr_all}. In addition, the less refined multi-stream grids are prone to noise at high $n_{str}$ as usual.
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig3a} 
  \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig3b}
\end{minipage}\hfill
\caption{Top: Volume fraction distribution of streams in 6 simulation boxes of size 100 $h^{-1}$ Mpc, 200 $h^{-1}$ Mpc 
and $128^{3}$, $256^{3}$, $512^{3}$ grids (with refinement factor of 1). Volume fractions are similar for simulation boxes with same inter-particle resolution. Slopes of the curve fits are shown in Table ~\ref{tab:Compare_Slopes}. Bottom: Volume fraction distribution for different refinement factors for 100 $h^{-1}$ Mpc, $128^{3}$ box. A considerably smoother volume fraction distribution is obtained at high number of streams in multi-stream fields with high refinement factor. }
\label{fig:Vfr_all}
\end{figure}

@@

Multi-stream field can be easily computed for a small Eulerian box with higher refinement factor. This can be utilized to analyse the phase-space behaviour inside and around haloes. In this section, 
we have used the halo coordinates identified by the FOF method, and selected Eulerian boxes around it. A reasonable  correspondence between FOF halo centres and local maxima of multi-stream field is visually examined in Fig. \ref{fig:full}.

@@

\caption{Multi-stream flow regions in a small box of the simulation. Top left: regions with more than 3-stream flow are identified as walls (brown). Intersection of multiple walls have higher $n_{str}$ regions (green, 17+ streams). Single-streaming voids (white) occupy large volume and are very close to the filaments in some directions. Top right: 17+ streams (green) form filamental structures with nodes at the intersections (red, 90+ streams)
Bottom left and right:  Closer look at the highly non-linear region reveals that a filament is sandwiched between the walls (brown). The 90+ stream region (red) forms a compact structure and is entirely contained within the filament. The black dots show the particles around the FOF halo within linking length of 0.2.}
\label{fig:4view}
\end{figure}

@@

Motivated by the complicated morphology of multi-stream field around a dark matter halo, we devised an empirical 
statistical tool to quantify the multi-stream environment of the FOF haloes. The method is geometrical and non-local.
We  randomly select a large number of points on a diagnostic spherical surface centred at the FOF centre of the halo 
and compute the number of streams  at these points. By increasing the radius of the sphere from inside of the halo to several times the halo radii, we estimate  the fractions of the area on the diagnostic spherical surface occupied by the regions with different numbers of streams:
3+, 5+, ..., where $n+$ corresponds to $n$ or higher number of streams.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig6a} 
  \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig6b}
\end{minipage}\hfill
\caption{Modelling a wall and a filament. A diagnostic spherical surface is intersected by a cylinder and plane.}
\label{fig:model}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig9a}
\includegraphics[width=10.cm]{Chapter3/Source_v2/fig9b} 
\end{minipage}\hfill
\caption{Halo of mass $7.0 \times 10^{12} M_{\odot} $  and  FOF radius 0.4 $h^{-1}$ Mpc. Top: All lines clearly scale between $r^{-1}$ and $r^{-2}$. Bottom: The filament is passing through the centre. It persists from radius of halo to 4 $h^{-1}$ Mpc. It is also surrounded by a single wall appearing as a line in the middle.}
\label{fig:706}
\end{figure}

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig10a}
\includegraphics[width=10.cm]{Chapter3/Source_v2/fig10b}
\includegraphics[width=6.cm]{Chapter3/Source_v2/fig10c}
\end{minipage}\hfill
\caption{Halo of mass $1.1 \times 10^{12} M_{\odot} $  and FOF radius 0.2 $h^{-1}$ Mpc (dotted-violet line). This halo just has 26 particles, hence the resolution is lesser than previous haloes for surfaces with low radii. Top: There is a bump in fraction of each of 3+ streams a little over 1 $h^{-1}$ Mpc. This is due to the presence of an additional halo nearby, as seen in the projections. The Mollweide projections from 0.4 $h^{-1}$ Mpc to 2 $h^{-1}$ Mpc have high stream flow regions near the lower surfaces of diagnostic spheres. Bottom: Corresponding FOF halo (red, at the centre) has a more massive neighbouring FOF halo (blue) within distance of 2 $h^{-1}$ Mpc. The 17+ stream regions (green) are increased around the neighbouring halo.} 
\label{fig:4322}
\end{figure}

@@

Up to 1 $h^{-1}$ Mpc from halo center of the largest halo, the surfaces are uniformly covered with high number of streams (red, 17+). This shows that the most non-linear regions are close to centres of haloes. A similar trend is seen for the halo of radius 0.7 $h^{-1}$ Mpc (Fig. ~\ref{fig:79}). However, for smaller haloes (Fig. \ref{fig:706} and \ref{fig:4322}) lower number of streams (even the wall forming 3+ streams; blue) start occupying the spherical surface at radii lesser than FOF-radius. In the case of the smallest halo of $1.1 \times 10^{12} M_{\odot} $ mass, the 17+ streams are seen at scales as low as 0.1 $h^{-1}$ Mpc. The distribution of multi-streams on the surface seems do not have a symmetry of any kind , signifying a complex morphology of the web in the vicinity of the haloes. Regions with 5+ to 15+ streams form structures intermediary to filament-like and wall-like behaviour, as seen by scaling of fraction of total points on the space with distance from halo center.

@@

Generally the transitions from haloes to filaments then to walls and finally to voids appear to be rather smooth.
However occasionally sharp features as the one seen in Fig. ~\ref{fig:4322} may emerge when the diagnostic sphere hits a neighbouring halo.

@@

\section{Summary}
In this chapter, we explore for the first time, the multi-stream environment of dark matter haloes in cosmological N-body simulations.
%We study the multi-stream field computed in N-body simulations in $\Lambda$CDM cosmology at $z=0$.
The visualization and quantitative characterization of generic non-linear fields in three-dimensional space represent a serious
challenge from both conceptual and computational points of view. The complexity of the problem requires diverse tools
for analysing the results of cosmological simulations as well as galaxy catalogues.

>>>>

The problem of objective identification and quantitative characterization of anisotropic structures 
in the distribution of galaxies in space emerged after the first evidences of their existence
(see the review by \citealt{Oort1983} and the references therein). The first theoretical model predicting highly anisotropic concentrations in the mass distribution coming into existence at the non-linear stage of gravitational instability is known as the Zel'dovich Approximation (the ZA) (\citealt{Zeldovich1970}, for further developments see also \citealt{Shandarin1989} and the references therein). The ZA predicted the formation of  `pancakes'  also known as the walls in the currently popular jargon. The later development
of the model by \citet{Arnold1982} predicted the formation of filaments along with the pancakes. \citet{Klypin1983a} and \citet{Shandarin1984} demonstrated that the filaments emerge in the cosmological N-body simulation in three-dimensional space.
However, they failed to identify the pancakes at $z=0$. Both the existence of filaments connecting compact clumps of matter and absence of pancakes were confirmed by \citet{Frenk1983}. Puzzled by the absence of the pancakes \citet{Klypin1983a}
speculated that insufficient mass resolution of the simulation was the cause  of the negative outcome. 
This has been unambiguously confirmed by recent simulations using a better numerical technique of computing a density field from the particle
coordinates in cosmological N-body simulations (~\citealt{Shandarin2012} and ~\citealt{Abel2012}). \citealt{Klypin1983a} 
also stressed that the most of filaments are  incorporated in 
`a single three-dimensional web structure'\footnote{The term `cosmic web' was coined by  \citet{Bond1996}.}. 
They admitted that
their simulation did not allow them to confirm the existence of pancakes between the filaments predicted by the ZA \citep{Arnold1982}.

Although the four archetypal elements of the cosmic web:  voids, walls/pancakes, filaments and haloes 
were predicted by ZA and confirmed in cosmological N-body simulation their identification and quantitative characterization
remains under vigorous debate
(see e.g. \citealt{Colberg2008}, \citealt{Elahi2013}, \citealt{Knebe2013}, \citealt{Hoffmann2014}).
The dark matter haloes are arguably the easiest objects to identify in N-body simulations. They can also be reliably associated with
observed objects like galaxies and clusters of galaxies.  But even in this case \citet{Knebe2013} refer to almost thirty different
halo finders suggested after 2000. Identifying filaments and pancakes/walls is far more controversial in both N-body simulations 
and galaxy catalogs. For instance, even estimating the global parameters of the  web  in N-body simulation such as the fractions of volume and mass
in voids, walls/pancakes, filaments and haloes produced quite different results. The estimates of volume 
fractions of voids range from 13  to 86\%  (\citealt{Cautun2014a}, \citealt{Falck2015}, \citealt{Forero-Romero2009a} , \citealt{Hahn2007} , \citealt{Aragon-Calvo2010a}). Similar estimates for walls/pancakes, filaments 
and haloes  are respectively 5-56\% 2-26\% and 0.1-1\%.  Estimates of the mass content vary in large ranges as well.

Large differences in the estimates of volume and mass fractions made by different groups are not surprising if we recognize
considerable differences in the definitions of the components of the cosmic web and numerical methods used in the estimates. Without trying to provide an exhaustive review of all definitions and techniques used for the quantitative morphological 
analysis of the web  we just briefly describe a few approaches in order to illustrate how different they could be. 
Some groups study the web morphology using only coordinates of simulation particles, while others use the particle velocities too. 
Transforming data from  point sets to the density and other fields on a grid is also often used
because fields allow to use a variety of mathematical techniques not available for the particle sets.
However this step can be done by a variety of methods some as simple as cloud-in-cell (CIC), or more complicated as smoothed-particle hydrodynamics (SPH), 
or using Voronoi and Delaunay tessellations
as in the Delaunay Tessellation Field Estimator (DTFE) method (~\citealt{Weygaert2009}, \citealt{Cautun2014a}).  
Recently a new method called a discrete persistent structure extractor (DisPerSE \citealt{Sousbie2011f},  \citealt{Sousbie2011e} allowing to identify haloes and other components of the web directly from the particles 
has been designed.  This method can be applied to the galaxy catalogs, for instance \citet{Sousbie2011e} applied it to SDSS catalog and extracted the filaments (which are made available online).

An obvious advantage of methods based on particle coordinates, both based on the density field and directly on particle coordinates, is their applicability to redshift catalogs. 
The redshift catalogs like SDSS and 2dF provide only two angular coordinates and distances in redshift space. 
But cosmological N-body dark matter simulations provide the full dynamical information in six-dimensional
phase space. This additional information is very valuable providing a greater opportunity for understanding the 
physics of the web and developing a better theory of the web.

@@

Using the full six-dimensional information allows one to generate new fields which  provide 
additional useful information about the evolution and morphology of the  web. 
One of them is a multi-stream field in Eulerian space, which will be the focus of this chapter. 
Another example is the flip-flop field in Lagrangian space. In cosmological context it was first used in the ZA.
\citet{Vogelsberger2011} used it in a study of multi-stream structure of galaxy size haloes.
\citet{Shandarin2014a} applied it for identifying sub-haloes  in dark matter haloes.
A similar although somewhat simplistic realization of this idea has been revealed in the ORIGAMI method
used for the analysis of the web ( \citealt{Falck2012},  \citealt{Falck2015} ).
 Although these fields cannot be used directly on observational data because the full phase-space information
 is not available, they provide much deeper insight into non-linear clustering of collision-less dark matter
and reveal new features of the web.

@@

In order to compute the multi-stream field we will use the tessellation scheme described in
 \citet{Shandarin2012} which is  also briefly discussed in Section \ref{subsec:method}.
  Using this methodology on the entire simulation box, we discuss the global behavior of the number of streams in the cosmic web in Section \ref{sec:global}. 
  The tessellation technique we have utilized can be used to find multi-stream fields in smaller Eulerian boxes with very high resolution too. In Section \ref{sec:local} we study the local behavior of multi-streams flows in regions around haloes the detected using friends-of-friends (FOF) technique.

@@

We have utilized the data from cosmological N-body simulations by Gadget-2 \citep{Springel2005c} for 100 $h^{-1}$ Mpc  and 200 $h^{-1}$ Mpc box 
sizes with $128^3$, $256^3$ and $512^3$ grids. Each particle is between  $10^9 - 10^{12} M_{\odot}$. 
The initial conditions and cosmological parameters are consistent with the Planck cosmology. We utilize the initial Lagrangian box and do a three-dimensional mapping onto corresponding evolved simulations. In addition, for local multi-stream analyses around haloes, we have utilized halo catalogs for each of these simulation boxes. These haloes are detected using FOF method considering objects with more than 20 particles found at linking length, $ b= 0.2$.  
%-------------------------------------------------------------------------------------------------------------------------------
\section{Multi-stream field calculation}
\label{subsec:method}

Phase space tessellation considers the dynamics of the particles similar to that of a standard N-body code. 
However the particles are nodes of the tessellation, and are just massless tracers of the flow.
Assuming that the uniform state is modeled by a simple rectangular grid, the particles are the nodes of the grid.
Each elementary cube of the grid is tessellated by five tetrahedra (\citealt{Shandarin2012}\footnote{For the description of an alternative type of the tessellation see \citet{Abel2012}.}) of which the vertices are the
vertices of the cube. Mass is assumed to be uniformly distributed within each tetrahedron and the tessellation remain
intact at all times. The tetrahedra of the tessellation change their shapes and volumes, the latter are used for
computing the densities of the tetrahedra. Despite the complicated deformations experienced by the three-dimensional
sub-manifold tessellated by the tetrahedra, it remains continuous. Projected on three-dimensional configuration space, the tetrahedra
may form complicated structures. The number of streams at a chosen point {\bf x} is simply the number of tetrahedra that contain the point. The diagnostic points are computationally convenient to choose on a regular grid which can be significantly finer than the original grid in Lagrangian space. The ratio of separation of particles on the initial unperturbed grid to the separation distance of points in the diagnostic grid  $l_{\rm part}/l_{\rm dg} $ will be referred to as the refinement factor in the rest of the chapter.

@@

Whereas for non-linear structures, our parameter space has more freedom in terms of number of streams. 
Similar to the density threshold, the number of streams - used as a local parameter - cannot distinguish unambiguously whether a point is in a wall, 
filament or halo. Only some parts of walls where there are only three streams can be identified locally without confusion. This is because the formation of a filament
requires at least five streams. A flip-flop along one axis  would produce a three-stream region which may be only a pancake. 
 Therefore  another flip-flop along the other axis in one of the streams
from previous stage is required to transform it into a three-stream flow. Thus the total becomes five. However, if the second flip-flop happens along the same axis 
the resultant structure will remain a wall. Therefore some points in the five-stream flows can be within walls while the other in filaments. The present simulations have no information about the evolution of the flip-flop field
therefore we rely on visual impressions initially to understand the transformation of walls into filaments and parts of filaments into  haloes. By inspection, we have identified all the regions with three streams as walls. Unfortunately, walls are difficult to display on paper since they essentially block the view in two-dimensional projection. Nevertheless, we have demonstrated and analyzed walls on a smaller Eulerian box around haloes in Section \ref{sec:local} using a simple and reasonably effective approximation.

@@

The filamentary structure of regions with 17 or more streams (denoted as 17+ ) is shown for a slice of simulation box of size $100 h^{-1}$  Mpc and $128^3$  particles in Fig. \ref{fig:full}. It has to be noted that all the regions with 17+ streams are regions with 3+ streams. Thus, the filaments are just interior
parts  of walls with higher $n_{str}$. These are visually observed mostly at the intersections of walls. Further, at the intersections of multiple filaments, there are regions with locally maximum number of streams, signifying the most dense regions in the simulations i.e. the dark matter haloes as Fig. \ref{fig:4view}
illustrates.
By superimposing the positions from the FOF-halo catalog, it is visually confirmed that the FOF haloes reasonably coincide with these high-streaming intersections, as Fig. \ref{fig:full} illustrates.

@@

One of the advantages of using tessellation is the freedom to compute densities at very high resolutions (\cite{Abel2012}, \cite{Shandarin2012}). We remind that the parameter `refinement factor' denotes the ratio of separation of the particles to the separation distance of points in the diagnostic grid as defined in Sec. 3. High refinement factors are extensively used in understanding stream behavior not only in the halo environment, but inside the halo too (Section ~\ref{sec:local}). The volume fractions of resulting number of streams are
%normalized by multi-stream field resolution (i.e. dividing by the cube of refinement factor) are %
similar for all refinement factors as
shown in bottom of the Fig. ~\ref{fig:Vfr_all} and in Table ~\ref{tab:Compare_ref}. % Hence our triangulation algorithm is universal in the sense that it is independent of the refinement factors used in tessellating the box.   
Multi-stream fields calculated on low refinement factors are more noisy at high number of streams.

With same refinement factors, the mass fractions exhibit similar pattern for same $L/N$ as well (Fig.~\ref{fig:Mfr_all}). The simulation box with highest inter-particle distance (thus least mass resolution) has more mass particles in single streaming region, as tabulated in Table~\ref{tab:Compare_LN}, but decreases steeply thereafter (Table ~\ref{tab:Compare_Slopes}). Unlike the volume fraction, the behavior of mass fraction has a systematic variation across different refinement factors. The mass fractions given in Table ~\ref{tab:Compare_ref} show that the single-streaming regions in the multi-stream fields with refinement factors 1 and 2 have higher mass fraction than in the fields with refinement factor of 8. Decreasing the resolution from refinement factor of 8 to 1 effectively introduces smoothing of the structure. This results in growth of mass fraction in voids ($n_{str} = 1$) and decreasing it in the web ($n_{str} > 1$). The multi-stream field is more robust as one can see in Fig.~\ref{fig:Vfr_all}. and \ref{fig:Mfr_all}. In addition, the less refined multi-stream grids are prone to noise at high $n_{str}$ as usual.
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig3a} 
  \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig3b}
\end{minipage}\hfill
\caption{Top: Volume fraction distribution of streams in 6 simulation boxes of size 100 $h^{-1}$ Mpc, 200 $h^{-1}$ Mpc 
and $128^{3}$, $256^{3}$, $512^{3}$ grids (with refinement factor of 1). Volume fractions are similar for simulation boxes with same inter-particle resolution. Slopes of the curve fits are shown in Table ~\ref{tab:Compare_Slopes}. Bottom: Volume fraction distribution for different refinement factors for 100 $h^{-1}$ Mpc, $128^{3}$ box. A considerably smoother volume fraction distribution is obtained at high number of streams in multi-stream fields with high refinement factor. }
\label{fig:Vfr_all}
\end{figure}

@@

Multi-stream field can be easily computed for a small Eulerian box with higher refinement factor. This can be utilized to analyze the phase-space behavior inside and around haloes. In this section, 
we have used the halo coordinates identified by the FOF method, and selected Eulerian boxes around it. A reasonable  correspondence between FOF halo centers and local maxima of multi-stream field is visually examined in Fig. \ref{fig:full}.

@@

\caption{Multi-stream flow regions in a small box of the simulation. Top left: regions with more than 3-stream flow are identified as walls (brown). Intersection of multiple walls have higher $n_{str}$ regions (green, 17+ streams). Single-streaming voids (white) occupy large volume and are very close to the filaments in some directions. Top right: 17+ streams (green) form filamentary structures with nodes at the intersections (red, 90+ streams)
Bottom left and right:  Closer look at the highly non-linear region reveals that a filament is sandwiched between the walls (brown). The 90+ stream region (red) forms a compact structure and is entirely contained within the filament. The black dots show the particles around the FOF halo within linking length of 0.2.}
\label{fig:4view}
\end{figure}

@@

Motivated by the complicated morphology of multi-stream field around a dark matter halo, we devised an empirical 
statistical tool to quantify the multi-stream environment of the FOF haloes. The method is geometrical and non-local.
We  randomly select a large number of points on a diagnostic spherical surface centered at the FOF center of the halo 
and compute the number of streams  at these points. By increasing the radius of the sphere from inside of the halo to several times the halo radii, we estimate  the fractions of the area on the diagnostic spherical surface occupied by the regions with different numbers of streams:
3+, 5+, ..., where $n+$ corresponds to $n$ or higher number of streams.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
 \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig6a} 
  \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig6b}
\end{minipage}\hfill
\caption{Modeling a wall and a filament. A diagnostic spherical surface is intersected by a cylinder and plane.}
\label{fig:model}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig9a}
\includegraphics[width=10.cm]{Chapter3/Source_v2/fig9b} 
\end{minipage}\hfill
\caption{Halo of mass $7.0 \times 10^{12} M_{\odot} $  and  FOF radius 0.4 $h^{-1}$ Mpc. Top: All lines clearly scale between $r^{-1}$ and $r^{-2}$. Bottom: The filament is passing through the center. It persists from radius of halo to 4 $h^{-1}$ Mpc. It is also surrounded by a single wall appearing as a line in the middle.}
\label{fig:706}
\end{figure}

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig10a}
\includegraphics[width=10.cm]{Chapter3/Source_v2/fig10b}
\includegraphics[width=6.cm]{Chapter3/Source_v2/fig10c}
\end{minipage}\hfill
\caption{Halo of mass $1.1 \times 10^{12} M_{\odot} $  and FOF radius 0.2 $h^{-1}$ Mpc (dotted-violet line). This halo just has 26 particles, hence the resolution is lesser than previous haloes for surfaces with low radii. Top: There is a bump in fraction of each of 3+ streams a little over 1 $h^{-1}$ Mpc. This is due to the presence of an additional halo nearby, as seen in the projections. The Mollweide projections from 0.4 $h^{-1}$ Mpc to 2 $h^{-1}$ Mpc have high stream flow regions near the lower surfaces of diagnostic spheres. Bottom: Corresponding FOF halo (red, at the center) has a more massive neighboring FOF halo (blue) within distance of 2 $h^{-1}$ Mpc. The 17+ stream regions (green) are increased around the neighboring halo.} 
\label{fig:4322}
\end{figure}

@@

Up to 1 $h^{-1}$ Mpc from halo center of the largest halo, the surfaces are uniformly covered with high number of streams (red, 17+). This shows that the most non-linear regions are close to centers of haloes. A similar trend is seen for the halo of radius 0.7 $h^{-1}$ Mpc (Fig. ~\ref{fig:79}). However, for smaller haloes (Fig. \ref{fig:706} and \ref{fig:4322}) lower number of streams (even the wall forming 3+ streams; blue) start occupying the spherical surface at radii lesser than FOF-radius. In the case of the smallest halo of $1.1 \times 10^{12} M_{\odot} $ mass, the 17+ streams are seen at scales as low as 0.1 $h^{-1}$ Mpc. The distribution of multi-streams on the surface seems do not have a symmetry of any kind , signifying a complex morphology of the web in the vicinity of the haloes. Regions with 5+ to 15+ streams form structures intermediary to filament-like and wall-like behavior, as seen by scaling of fraction of total points on the space with distance from halo center.

@@

Generally the transitions from haloes to filaments then to walls and finally to voids appear to be rather smooth.
However occasionally sharp features as the one seen in Fig. ~\ref{fig:4322} may emerge when the diagnostic sphere hits a neighboring halo.

@@

\section{Summary}
In this chapter, we explore for the first time, the multi-stream environment of dark matter haloes in cosmological N-body simulations.
%We study the multi-stream field computed in N-body simulations in $\Lambda$CDM cosmology at $z=0$.
The visualization and quantitative characterization of generic non-linear fields in three-dimensional space represent a serious
challenge from both conceptual and computational points of view. The complexity of the problem requires diverse tools
for analyzing the results of cosmological simulations as well as galaxy catalogs.
